%!TEX root=cs440S2014-lab7.tex 
% mainfile: cs440S2014-lab7.tex 

\input{labspre.tex}

\usepackage[compact]{titlesec}

\begin{document} \MYTITLE{Laboratory Assignment Seven: Evaluating the Efficiency of File System Searching}
\MYHEADERS{Laboratory Assignment Seven}{Due: April 7, 2014}

% \vspace*{-.32in}

\section*{Introduction}

Since most software applications ultimately need to store and retrieve information, it is essential that an operating
system  provide a file system to support these activities.  It is also important to ensure that application programs
effectively use the primitives provided by the file system.  In this laboratory assignment, we will explore and
experimentally evaluate two file system searching tools, {\tt grep} and {\tt ag}. One of the tools that we will study,
{\tt grep}, was first implemented years ago by Ken Thompson, one of the creators of Unix.  The second tool, known as
{\tt ag} or the ``Silver Searcher'', is a new tool for file system searching. As part of this assignment, you will learn
the command lines associated with both of these tools and then design a benchmarking framework to determine which one is
faster.  Finally, you will explore at least one advanced application of these tools.

\section*{File System Searching with {\tt grep}, {\tt ag}, and {\tt nautilus}}

Before you start using these two tools, you should learn about their history. For instance, you should learn exactly
when the tools were implemented and who initially created them. Next, you can run the {\tt man} command in your terminal
window and study its output to ensure that you understand the command lines that both tools support. What are the basic
commands for searching a specific directory with these two tools? Finally, you should type {\tt nautilus} in your
terminal window to run Ubuntu's graphical file system browser.  How does {\tt nautilus} support file system searching?
Overall, which tool(s) seem to furnish the most robust set of searching features? Why?

\section*{Experimentally Evaluating File System Searchers}

When you are implementing a large software system, like an operating system or a compiler, it is important for you to have
the ability to quickly search through a large collection of files for those that contain a specified word or pattern.
For this part of the laboratory assignment, you should find five separate directories of source code that contain a
substantial number of files. Next, you should use a Linux command to determine how many files are in each of the chosen
directories; please make sure that you clearly explain how you counted the total number of files.

The next step in the experimental evaluation of {\tt grep} and {\tt ag} is to devise at least five different command
lines that you can run for each of the tools. Then, for each of the five commands and the five different directories
containing code, you should evaluate the efficiency and effectiveness of the tools. In particular, you should evaluate
the searchers according to the following metrics: (i) the number of results produced, (ii) the usefulness and accuracy
of the output, and (iii) the speed at which the tools produce their output. When conducting the experiments to evaluate
efficiency, please run multiple trials and report arithmetic means and standard deviations. After you have completed
this basic experiment, you should develop and run one additional study. For instance, you could compare the performance
of file system searchers that are run in both your home account (e.g., {\tt /home/g/gkapfham/}) and on a local disk
(e.g., the {\tt /tmp} directory on your workstation).

\section*{Analysis of the Searching Results}

Now that you have defined your benchmarks and run your experiments, you should analyze the searching results. Beyond
identifying which tool is faster and the circumstances in which it is better, you should also aim to explain, whenever
it is possible to do so, why these results are evident. If you had to pick one of these tools for searching both large
and small directories of source code, which one would you select? Why? As you complete a report on the results from your
empirical study, make sure that you give all of the relevant details that would be needed to support the replication of
your results. Finally, you should comment on whether or not it would be possible to perform the same type of file
searching with {\tt nautilus}, Ubuntu's graphical file system browser. Please see the instructor if you have questions
about completing these analysis tasks.

\section*{Advanced Use of File System Searchers}

File system searching tools like {\tt grep} and {\tt ag} have a wide variety of interesting applications.  For instance,
the Vim text editor can be extended with CTRL-P, a plugin that supports fuzzy file finding.  You can learn more about
CTRL-P by visiting the following Web site: \url{http://kien.github.io/ctrlp.vim/}. One critique of CTRL-P is that it is
only fast enough for regular use when it leverages caching.  However, users will often invalidate the CTRL-P cache when
they create new files, forcing the cache to be rebuilt with a time consuming additional scan. One resolution to this
problem is to configure CTRL-P to use an external user command such as {\tt ag}.  As such, an example of an advanced use
of file system searchers that you could report on is the configuration of Vim to use an enhanced version of CTRL-P that
leverages {\tt ag}. Alternatively, you could install and explain how to use the {\tt ag.vim} plugin that makes it
possible to run {\tt ag} searches directly from Vim. Since Vim also provides integration with the {\tt grep} command,
you could explore this advanced feature as well. Students may also look into advanced features that do not involve the
use of Vim.

\section*{Summary of the Required Deliverables}

This assignment invites you to submit printed and signed versions of the following deliverables: 

\vspace*{-.05in}
\begin{enumerate}

  \itemsep0in

  \item A description of the history, features, and drawbacks associated with {\tt grep}, {\tt ag}, and {\tt nautilus} 

  \item The command lines that you used to run all of your benchmarks with {\tt grep} and {\tt ag} 

  \item Screenshots demonstrating that you could correctly configure and run {\tt grep}, {\tt ag}, and {\tt nautilus} 

  \item A comprehensive report that explains the results and responds to all of the stated questions 

  \item The source code and/or documentation for an advanced use of the file system searching tools

  \item Ideas for future experiments and analyses that use and evaluate the file system searchers 

  \item A reflective discussion of the challenges that you encountered when completing this assignment

\end{enumerate}
\vspace*{-.05in}

In adherence to the honor code, students should complete this assignment on an individual basis. While it is appropriate
for students in this class to have high-level conversations about the assignment, it is necessary to distinguish
carefully between the student who discusses the principles underlying a problem with others and the student who produces
assignments that are identical to, or merely variations on, someone else's work.  As such, deliverables that are nearly
identical to the work of others will be taken as evidence of violating the \mbox{Honor Code}.  

  \end{document}
